### 2016-07-12 30June16_plate_3_explant_sup.Rmd
First commit

---
title: '30June16 explant sups plate 3: MIG, RANTES, IL1ra'
author: "Claire Levy"
date: "July 6, 2016"
output: github_document

---
This is an analysis of data from running our MSD plate 3 on some of HSV2 infected explant sup samples for Herpes P01. See explant_and_HVE_cell_sup_MSD_README for more on the different plates/analyte panels. This plate has the following analytes measured in each well

Plate 3: 96 well 4 spot prototype Human Triplex

	Il-1ra
	MIG (CXCL9)
	RANTES (CCL5)

The MSD machine gives the data in an excel or text format. I have .csv file with the results.

```{r setup, echo=FALSE, warning=FALSE, message= FALSE}
library(dplyr)
library(ggplot2)
library(pander)
library(reshape2)
library(stringr)
library(swarm)

#read in the .csv raw results file
#../ to go up one level and then back down into a different folder. I exported this file from the DWB experiment called "Hladik plate 3 30 June 16 corrected layout". I imported that file from the folder location where FC saved the DWB experiment with the corrected layout. That file is here:

#J:\MacLabUsers\HLADIK SHARED\Projects\Herpes study\MSD\CVL Test Samples\Raw Data from MSD\Test Run

#I appended the name with "corrected layout" because I imported the wrong file at first and you can't delete them from the software.

dat<- read.csv("../raw_data/30June16_plate3_CVL_and_explant_sup_CLexport.csv", skip = 1)

#### NOTE ####
# This file also contains data from TEST CVL samples from the hvtn clinic that we were running to test the analytes before doing the actual trial samples from Kenya.

#make a column identifying what group the sample belongs to, standards or explant sup or CVL sample so I can filter out what I want.

dat<- dat %>%
  mutate(Type =
 ifelse(str_detect(Sample,"^3")==TRUE, "explant_sup",
        ifelse(str_detect(Sample, "R|V")==TRUE,
                          "CVL","Standard")))

## Also, we included the diluted stock calibrator and the blended working stock on the plate when we shouldn't have. So we excluded those samples in the Discovery workbench software, which re-sets the curve fit based on the remaining samples. When you export the data table, it still includes those EXCLUDED samples. I am removing them here.

dat <- dat %>%
  filter(Sample != "Diluted Stock Calibrator" & Sample != "Blended Working Stock")
```

```{r standards, echo=FALSE, warning=FALSE, message= FALSE}

#filtering out the standards
standards <- dat %>%
  filter(Type == "Standard")

#which standards are in the detection range
standardsInRange <-standards%>%
  filter(Detection.Range == "In Detection Range")

#which were not in the detection range?

standardsOutOfRange<-standards%>%
  filter(Detection.Range != "In Detection Range")%>%
  arrange(Type, Assay, Sample)%>%
  select(Type,Sample, Assay, Detection.Range, X..Recovery)

```

Of the `r length(standards)` standards that we measured, `r length(standardsInRange)` are in the detection range. The following standards were out of the detection range:
`r pander(standardsOutOfRange)`

Here is a plot of the % recovery (extrapolated concentration/ known concentration x 100) for the standards that were in the detection range, with lines at the "good recovery" limits of 80 and 120% recovery.

```{r percent recovery, echo=FALSE, warning=FALSE, message= FALSE}

stdFit<- standardsInRange %>% 
  arrange(Sample,Assay)%>%
  select(Type, Sample, Assay, X..Recovery,Fit.Statistic..RSquared, Concentration, Calc..Concentration)

ggplot(stdFit, aes(x = Sample, y = X..Recovery))+
  geom_point(aes(color = Sample))+
  facet_wrap(~Assay)+
  theme(axis.text.x = element_text(angle=25))+
  scale_y_continuous(limits = c(75, 200))+
  geom_abline(slope = 0, intercept = c(80, 120))+
  labs(y = "Percent Recovery")

```

The R-squared values for goodness-of-fit were good for all assays:

```{r r-squared, echo=FALSE, warning=FALSE, message= FALSE}

rSquared<-stdFit%>%
  group_by(Assay)%>%
  summarize(R_Squared = mean(Fit.Statistic..RSquared))

pander(rSquared)
```

```{r explant data cleanup, echo=FALSE, warning=FALSE, message= FALSE}

#filtering for the samples I want (just want explant sups, not CVLS or standards)

explantDat<-dat%>%
  filter(Type == "explant_sup")

#sanity check: I put 56 samples on the plate and there are 3 analytes so we should have 56 x 3 = 168 samples.

#We didn't have enough sample to do two replicates for 324-T2-M so we didn't add anything to well F10. We did read the well though, so there is data for it, but I want to exclude it because there wasn't sample there

explantDat <- explantDat %>%
  filter(Well != "F10") #removes 3 data points, 1 for each analyte so 165 left

#which explant samples are in the detection range
explantDatInRange <-explantDat%>%
  filter(Detection.Range == "In Detection Range")

#which were not in the detection range?

explantDatOutOfRange<-explantDat%>%
  filter(Detection.Range != "In Detection Range")%>%
  arrange(Type, Assay, Sample)%>%
  select(Type,Sample, Assay, Detection.Range)

```

```{r adding ids and avg reps, echo=FALSE, warning=FALSE, message= FALSE}

# I want to be able to group the data by timepoint and donor and virus so I need to separate the sample name and make new columns identifying them that way

donorTimeVirus<-colsplit(explantDat$Sample,"-", c("Donor","Timepoint","Virus") )

#bind the split col back to explantDat

explantDat<-cbind(donorTimeVirus, explantDat)

#I should only have 1 replicate for 324-T2-M. All other samples have 2 reps.

#There is one sample where the concentration was below the fit curve range and the concentration is given as "NaN". In order for that point to appear on the plot, I am going to give those samples a value of "1"

NaNSamples<- explantDat %>%
  filter(Calc..Concentration == "NaN")%>%
  select(Sample, Detection.Range, Calc..Concentration)

explantDat <- explantDat %>%
  mutate(Calc..Concentration = ifelse(Calc..Concentration == "NaN", "1", Calc..Concentration))

explantDat$Donor <- as.factor(explantDat$Donor)

explantDat$Calc..Concentration <-as.numeric(explantDat$Calc..Concentration)

explantDat$Type <- as.factor(explantDat$Type)

explantDat$Virus <- as.factor(explantDat$Virus)

explantDat$Virus<-factor(explantDat$Virus, levels = c("M", "V1","V2","M 1:50", "V2 1:50"))

explantDat$Timepoint<-as.factor(explantDat$Timepoint)

#Averaging the replicates. 
#Sanity check: 
#For 27/28 samples I have 2 replicates = 54 
# For 1/28 samples I have 1 replicate = 1
#If I average those samples, I get 54/2 + 1 = 28 avg values
# I have data for 3 analytes so I should have 28 * 3 average values.

AvgConcExplantDat <- explantDat %>%
  group_by(Sample, Assay, Donor, Timepoint,Virus, Detection.Range)%>%
  summarize(avgConc = mean(Calc..Concentration))
```

MIG data

```{r MIG, echo=FALSE, warning=FALSE, message= FALSE}

#I want to make separate plots for each analyte, so I'll split it into different dfs here:

#MIG

MIGAvgConcExplantDat<- AvgConcExplantDat %>%
  filter(Assay == "MIG")

ggplot(data = MIGAvgConcExplantDat, aes(x = Virus, y = avgConc))+

  geom_point(aes(color = Donor, shape = Detection.Range ),size = 2 )+

  geom_line(aes(color = Donor, group = Donor))+

  scale_shape_manual(values = c("Above Fit Curve Range" = 1, "Below Fit Curve Range" = 1, "In Detection Range" = 16))+

  scale_y_log10()+

  facet_wrap(~Timepoint)+

  theme(axis.text.x = element_text(angle = 25))
```

RANTES data

```{r RANTES, echo=FALSE, warning=FALSE, message= FALSE}
#RANTES

RANTESAvgConcExplantDat<- AvgConcExplantDat %>%
  filter(Assay == "RANTES")

ggplot(data = RANTESAvgConcExplantDat, aes(x = Virus, y = avgConc))+

  geom_point(aes(color = Donor, shape = Detection.Range ),size = 2 )+

  geom_line(aes(color = Donor, group = Donor))+

  scale_shape_manual(values = c("Above Fit Curve Range" = 1, "Below Fit Curve Range" = 1, "In Detection Range" = 16))+

  scale_y_log10()+

  facet_wrap(~Timepoint)+

  theme(axis.text.x = element_text(angle = 25))
```

IL1-RA data

```{r IL1-ra, echo=FALSE, warning=FALSE, message= FALSE}
## IL1-ra

IL1raAvgConcExplantDat<- AvgConcExplantDat %>%
  filter(Assay == "IL1-RA")

ggplot(data = IL1raAvgConcExplantDat, aes(x = Virus, y = avgConc))+

  geom_point(aes(color = Donor, shape = Detection.Range ),size = 2 )+

  geom_line(aes(color = Donor, group = Donor))+

  scale_shape_manual(values = c("Above Fit Curve Range" = 1, "Below Fit Curve Range" = 1, "In Detection Range" = 16))+

  scale_y_log10()+

  facet_wrap(~Timepoint)+

  theme(axis.text.x = element_text(angle = 25))

```