### 2016-08-23 MSD_panels_1_2_3_HVE_sup_analysis.Rmd
Changed the layout of the individual plots

---
title: "MSD panels 1, 2 and 3 HVE Sup"
author: "Claire Levy"
date: "August 22, 2016"
output: github_document
---

## Experiment overview

These data are from running MSD analysis of supernatents from selected ex vivo vaginal epithelial cells from three donors. The explants were infected with HSV2 strains 186, SD90 or Mock infected by Lamar Fleming and Gabriella Fenkart and supernatents were collected at 3 timepoints post-infection: 3hrs, 8hrs and 24hrs.

We used 3 different MSD panels containing a total of 20 analytes analyzed for 24 samples. Based on results from running test samples and from running the explant supernatent samples, we decided to dilute all the supernants 1:2 (plate 1) and 1:5 (plate 2) with MSD's diluent 43 before running the plates. All samples were run in duplicate.

```{r read in data, message=FALSE, warning = FALSE, echo = FALSE}

library(dplyr)
library(ggplot2)
library(pander)
library(reshape2)
library(stringr)
library(swarm)

#panel 1
# 128 wells x 9 analytes = 1152 obs
HVE_panel_1 <- read.csv("../raw_data/17Aug16_HVE_sup_panel_1_plate_1_and_2.csv", skip = 1)

#panel 2
#128 wells x 8 analytes = 1024 obs
HVE_panel_2 <- read.csv("../raw_data/17Aug16_HVE_sup_panel_2_plate_1_and_2.csv", skip = 1)

#panel 3
#128 wells x 3 analytes = 384 obs

HVE_panel_3 <- read.csv("../raw_data/18Aug16_HVE_sup_panel_3_plate_1_and_2.csv", skip = 1)

#combine data from all 3 panels

HVE_all_panels <-rbind(HVE_panel_1, HVE_panel_2, HVE_panel_3)

#Fix the greek

HVE_all_panels$Assay <-recode(HVE_all_panels$Assay,
                                      "IFN-α2a" = paste("IFN-\U03B1","2a", sep = ""),
                                      "IL-1α" = "IL-1\U03B1",
                                      "MIP-1α"="MIP-1\U03B1",
                                      "TNF-α" = "TNF-\U03B1",
                                      "MIP-3α" = "MIP-3\U03B1",
                                      "MIP-1β" = "MIP-1\U03B2",
                                      "IL-1β" = "IL-1\U03B2",
                                      "IFN-γ" = "IFNg")

#Just keep the columns that I care about

HVE_all_panels <- HVE_all_panels %>%
  select(Assay,Sample, Sample.Group, Plate.Name, X..Recovery, Concentration, Calc..Concentration, Detection.Range, Fit.Statistic..RSquared,Spot)

```

## Checking Standard Recovery and Curves

Here are plots of the % recovery (extrapolated concentration/ known concentration x 100) for the standards that were either within or above the detection range, with lines at the "good recovery" limits of 80 and 120% recovery.

The R^2^ values for goodness-of-fit for the standard curves were all >0.99

```{r standards, echo=FALSE, message = FALSE, warning = FALSE}

HVE_stds_all_panels<-filter(HVE_all_panels, Sample.Group == "Standards")

#which standards are out of det. range?
standardsOutOfRange<-HVE_stds_all_panels %>%
  filter(Detection.Range != "In Detection Range")%>%
  select(Plate.Name,Sample, Assay, Detection.Range, X..Recovery, Calc..Concentration)

#looking at which samples have STDs below the fit curve or below the detection range (i.e. you can't get numbers for % recovery)

HVE_stds_all_panels_below<- HVE_stds_all_panels%>%
  filter(Detection.Range == "Below Fit Curve Range" | Detection.Range == "Below Detection Range") %>%
  group_by(Sample)%>%
  summarise("Number of Samples Below Fit Curve Range/Detection Range" = n())

#The samples that actually give me #'s for percent recovery ( i.e. all but the "below fit curve range" and "below detection range" ones)

HVE_stds_all_panels_percent_recov<- HVE_stds_all_panels %>% 
  arrange(Sample,Assay)%>%
  filter(Detection.Range != "Below Fit Curve Range" & Detection.Range!= "Below Detection Range")%>%
  select(Plate.Name,Sample, Assay,Detection.Range, X..Recovery,Fit.Statistic..RSquared, Concentration, Calc..Concentration)

#plotting percent recovery

ggplot(HVE_stds_all_panels_percent_recov,aes( x = Sample, y = X..Recovery)) +
  geom_point(aes(shape = Detection.Range ),
             size = 2, alpha = 0.5)+
  scale_shape_manual(values = c("Above Fit Curve Range" = 1, 
                                "In Detection Range" = 16, 
                                "Below Detection Range" = 0))+
  facet_wrap(~Assay)+
  theme(axis.text.x = element_text(angle=25))+
  scale_y_continuous(limits = c(0, 265))+
  geom_abline(slope = 0, intercept = c(80, 120))+
  labs(y = "Percent Recovery")+
  theme(axis.text.x = element_text(size = rel(0.8)))

#checking the R squared goodness-of-fit for the standard curves
rSquared<-HVE_stds_all_panels_percent_recov%>%
  group_by(Assay)%>%
  summarize(R_Squared = mean(Fit.Statistic..RSquared))
```

## Caveats about Standards

* Standards with values *below the detection range* (<2.5x the standard deviation of the lowest standard) give unreliable concentrations, so I did not include those in the plots.

* When standards are *below the curve fit* (outside the range of the standard curve), the software does not calculate concentrations so you can't get a percent recovery.

* Most of the Standards that fell into either of these categories were from STD-07 or STD-08 except for five that were STD-06:

`r pander(HVE_stds_all_panels_below)`

```{r break up sample name, echo=FALSE, message = FALSE, warning = FALSE}
#Making a df that contains just the non-standards

HVE_all_panels_noStds <- HVE_all_panels %>%
  filter(Sample.Group!="Standards")

# I want to be able to group the data by timepoint and donor and virus so I need to break up components of the the sample name and make new columns with just the Donor, Timepoint, Virus and Sample_Dilution.

#Split samples names on the "-"
donorTimeVirusDil<-colsplit(HVE_all_panels_noStds$Sample,"-", c("Donor","Timepoint","Virus", "Sample_Dilution") )

#bind the split col back to allPanels

HVE_all_panels_noStds<-cbind(donorTimeVirusDil, HVE_all_panels_noStds)

#Make things factors so they plot nicely
HVE_all_panels_noStds$Donor <- as.factor(HVE_all_panels_noStds$Donor)

HVE_all_panels_noStds$Calc..Concentration <-as.numeric(HVE_all_panels_noStds$Calc..Concentration)

HVE_all_panels_noStds$Virus <- as.factor(HVE_all_panels_noStds$Virus)

HVE_all_panels_noStds$Virus<-factor(HVE_all_panels_noStds$Virus, levels = c("M", "V1","V2"))

HVE_all_panels_noStds$Timepoint<-as.factor(HVE_all_panels_noStds$Timepoint)

HVE_all_panels_noStds$Sample_Dilution<-as.factor(HVE_all_panels_noStds$Sample_Dilution)
```

## HVE Supernatents: Detection Range

The following plots show that the 1:2 dilution of the samples resulted in more samples being in "In Detection Range."

```{r checking detection range, echo=FALSE, message = FALSE, warning = FALSE}

#What number of samples are in each category of detection? Also adding a column to differentiate which analyte is in which panel, mostly for facetting reasons.

HVE_all_panels_noStds_NumInEach <- HVE_all_panels_noStds%>%
  mutate(Panel = ifelse(str_detect(Plate.Name, "panel_1"),"Panel 1",ifelse(str_detect(Plate.Name,"panel_2"),"Panel 2","Panel 3")))%>%
 group_by(Plate.Name,Sample_Dilution,Panel,Assay, Detection.Range)%>%
  summarise(Number_of_Samples = n())

#Let's plot that

ggplot(HVE_all_panels_noStds_NumInEach, aes(x = Assay, y=Number_of_Samples))+
         geom_point(aes(color = Detection.Range),size = 2)+
  theme(axis.text.x = element_text(angle=25, size = 8))+
  labs(y = "Number of Samples", x = "Analyte")+
  facet_wrap(Sample_Dilution~Panel, scales = "free")+
  ggtitle("Number of Samples in each Detection Range per Analyte")
```

```{r multiplying by dilution factor, echo=FALSE, message = FALSE, warning = FALSE}

#I'm going to make a column for estimated stock concentration by multiplying the avgConc by 2 or 5 (since I did a 1:2 and 1:5 dilution on the samples) so I can better compare with the explant sup concentrations.

HVE_all_panels_noStds <- HVE_all_panels_noStds%>%
  mutate(Estimated_Stock_Conc = ifelse(
    Sample_Dilution == "1:2", Calc..Concentration * 2,
    ifelse(Sample_Dilution == "1:5", Calc..Concentration *5, NA)
  ))

```

```{r exclude belows, echo=FALSE, message = FALSE, warning = FALSE}

#I am excluding samples that are "Below Detection Range" because this means that the signal was "statistically indistinguishable from the background or top of the curve"" and ..."The concentrations calculated are inherently unreliable." If one of two reps fell into either of these categories, the remaining replicate will serve as the "average" for that sample.

HVE_all_panels_noStds_noBelow <- HVE_all_panels_noStds %>%
  filter(Detection.Range != "Below Fit Curve Range" & Detection.Range != "Below Detection Range")
```

## All Samples Estimated Stock Concentrations

* I multiplied by the dilution factors so this is an estimated concentration of the stock samples.

* Note that this is on a log(10) scale; most concentrations are a lot lower than the high ones.

```{r  all sample concentrations, echo=FALSE, message = FALSE, warning = FALSE}

#Plot of all samples and their concentrations

ggplot(HVE_all_panels_noStds_noBelow,aes( x = Timepoint, y = Estimated_Stock_Conc))+
         geom_point(aes(color = Assay),position = position_swarm (0.5),size = 2)+
  scale_y_log10()+
  labs(y = "Estimated Stock Concentration pg/ml")+
  facet_wrap(~Virus)

```

```{r filter for 1:2, echo=FALSE, message = FALSE, warning = FALSE}
### NOTE ### I'm going to filter for just the 1:2 dilution samples here since they gave more in-detection.

HVE_all_panels_noStds_1to2_noBelow<-HVE_all_panels_noStds_noBelow %>%
  filter(Sample_Dilution =="1:2")

```

```{r avg reps, echo=FALSE, message = FALSE, warning = FALSE}

#averaging the replicates
avgConc_HVE_all_panels_noStds_1to2_noBelow <- HVE_all_panels_noStds_1to2_noBelow %>%
group_by(Sample, Assay, Donor, Timepoint,Virus)%>%
summarize(avgEstConc = mean(Estimated_Stock_Conc))%>%
  ungroup()

#Looking at the number of samples in each sample "set"(Same Donor and Timepoint). Some are missing V1 or V2. or both!! I don't want the ones with just one sample in the set.

numberInSets<-avgConc_HVE_all_panels_noStds_1to2_noBelow %>%
         group_by(Donor,Timepoint,Assay)%>%
  select(Sample,Donor, Timepoint,Assay)%>%
  summarize(numInSet = n())

#now I'll merge the avgConc df with this numberInSets info created above so I can then filter out samples that have <2 samples in the set

avgConc_HVE_all_panels_noStds_1to2_noBelow <- merge(avgConc_HVE_all_panels_noStds_1to2_noBelow, numberInSets, by = c("Donor", "Timepoint", "Assay"))

#now exclude the samples where there not a "full set", i.e. 1 sample only in the set

avgConc_HVE_all_panels_noStds_1to2_noBelow_fullSet <- avgConc_HVE_all_panels_noStds_1to2_noBelow %>%
  filter(numInSet != 1)

```

## Plots of analyte concentrations
Since the 1:2 diluted samples were more likely to be in the detection range, I am only showing data for those.

## Caveats
Samples omitted from the plots include:

* Concentrations below the fit curve range. Values cannot be extrapolated when the data was *below* the fit curve range. However,the MSD software can extrapolate values that are *above* the fit curve range if the curve is linear at the top.

* Samples for which there was only data for a Mock sample and neither of the virus conditions.

Also note:
If a sample had one rep that was below detection and another that was in range, I used the remaining "in range" rep as the "average" for that sample.

```{r  individual analyte plots , echo=FALSE, message = FALSE, warning = FALSE}

ggplot(avgConc_HVE_all_panels_noStds_1to2_noBelow_fullSet, aes(x = Timepoint, y = avgEstConc))+
   geom_point(aes(color = Donor), alpha = 0.75, size = 2)+
  scale_y_log10()+
   labs( y = "Estimated Stock Concentration pg/ml")+
   facet_grid(Assay~Virus, scales = "free_x")+
   theme(axis.text.x = element_text(size = 11))

################## another way of making the plots #########

#make a list of the analytes

#I'm using a unique vector here instead of the levels of Assay because I don't want to include IFNa2a OR Il-7. There weren't any data points in the detection range for those analytes (so I don't want to try to make a plot of it)

# analyteList<-as.list(unique(avgConc_HVE_all_panels_noStds_1to2_noBelow_fullSet$Assay))

# #write a function to make a plot for each analyte
# 
# plotAnalytes <- function (Analyte){
#    avgConc_HVE_all_panels_noStds_1to2_noBelow_fullSet %>%
#      filter(Assay == Analyte)%>%
#      
#      ggplot( aes(x = Timepoint, y = avgEstConc))+
#    geom_point(aes(color = Donor ), alpha = 0.5, size = 3)+
#   geom_line(aes(group = Donor, color = Donor), alpha = 0.5)+
#     stat_summary(fun.y = mean, geom = "point")+
#     stat_summary(fun.y = mean, geom = "line",aes(group = 1))+
#    scale_shape_manual(values = c("Above Fit Curve Range" = 1, "Below Detection Range" = 1, "In Detection Range" = 16))+
#    scale_y_log10()+
#    labs( y = "Concentration pg/ml")+
#   facet_wrap(~Virus, scales = "free_x")+
#    theme(axis.text.x = element_text(size = 11))+
#      ggtitle(Analyte)
# }
# 
# 
# 
# lapply(analyteList, FUN = plotAnalytes)

```