### 2016-08-23 MSD_panels_1_2_3_explant_sup_analysis.Rmd

---
title: "MSD panels 1, 2 and 3 Explant Sup"
author: "Claire Levy"
date: "August 12, 2016"
output: github_document
---

## Experiment overview

These data are from running MSD analysis of supernatents from selected vaginal explant samples from seven donors. The explants were infected with HSV2 strains 186, SD90 or Mock infected by Lamar Fleming and Gabriella Fenkart and supernatents were collected at 3 timepoints post infection: 3hrs, 8hrs and 24hrs.

We used 3 different MSD panels containing a total of 20 analytes analyzed for 46 samples (see caveat). Based on results from running test samples, we decided to dilute all the supernants 1:5 with MSD's diluent 43 before running the plates. All samples (except 1, see caveats) were run in duplicate.

## Caveats

Due to limited sample volume, we had to make the following omissions:

Panel 2: Only 1 rep for sample 324-T2-V1

Panel 3: Did not run samples 324-T2-V1 or 324-T2-V2.

Due to a possible pipetting error, we were not able to use the data from the standards that we ran for panel 1. Instead, we used the standards from a different run of the same panel to calculate concentrations for the experimental samples (see explant_sup_README for details).

```{r data read-in, echo=FALSE, message = FALSE, warning = FALSE}

library(dplyr)
library(ggplot2)
library(pander)
library(reshape2)
library(stringr)
library(swarm)

#reading in data from the panel 1

#remember that this data table contains data from the 17Aug16_HVE_sup_panel_1_plate_1. I am only going to use the Standards data from that plate and not the standards from either of the plates for panel 1 run 10Aug16 since we had what appears to be a pipetting error (2nd replicate is more concentration for all wells).

explant_panel_1_and_HVE_panel_1_plate_1 <- read.csv("../raw_data/10Aug16_explant_sup_panel_1_plate_1_and_2.csv", skip = 1)

#Filter to ONLY keep the standards from the plate called "HVE_sup_panel_1_plate_1"

StdsToUse_explant_panel_1 <- explant_panel_1_and_HVE_panel_1_plate_1 %>%
  filter(Sample.Group == "Standards")%>%
  filter(Plate.Name == "HVE_sup_panel_1_plate_1")

#Since I told Discovery Workbench software that I wanted to use the standards from the 17Aug16_HVE_sup_panel_1_plate_1 as "Global Standards" (i.e. use them to calculate concentrations for all plates in the experiment)it seems to have provided me with 3 copies of those samples in the data table (1 for each plate I guess). I only need data from the unique entries, so I will filter for just unique rows in this df to exclude copies.

# should have 8 stds x 9 analytes x 2 reps = 144 obs
StdsToUse_explant_panel_1<-unique(StdsToUse_explant_panel_1)

#Sanity check: there should only be 2 entries for each unique instance of Sample (STD-0x) and assay. 
checkStdReps<-StdsToUse_explant_panel_1 %>%
         group_by(Sample,Assay)%>%
         summarise(n())

#Remove the 17Aug16_HVE_sup_panel_1_plate_1 non-STD data from the df and leave ONLY explant panel 1 data. Remaining should be:
#94 wells from plate 1 + 30 wells from plate 2 x 9 analytes = 1116 observations

explant_panel_1 <-explant_panel_1_and_HVE_panel_1_plate_1 %>%
  filter(Plate.Name != "HVE_sup_panel_1_plate_1")

#now remove the unwanted Std samples from the explant data. There should be 32 Std wells x 9 analytes = 288
# 1116 total - 288 = 828 observations

No_Stds_explant_panel_1 <-explant_panel_1 %>%
  filter(Sample.Group != "Standards")

#now add in the unique Stds from 17Aug16_HVE_sup_panel_1_plate_1

corrected_Stds_explant_panel_1 <- rbind(No_Stds_explant_panel_1,StdsToUse_explant_panel_1)

# Data from panels 2 and 3

#panel 2 should have 123 wells (left out one rep) x 8 analytes = 984
explant_panel_2<-read.csv("../raw_data/10Aug16_explant_sup_panel_2_plate_1_and_2.csv", skip = 1)

#panel 3: 120 wells (left out 2 samples, 4 reps) x 8 analytes =360

explant_panel_3<-read.csv("../raw_data/11Aug16_explant_sup_panel_3_plate_1_and_2.csv", skip = 1)

#Combining the data from all 3 panels: 972 +984 + 360= 2316

explant_all_panels<- rbind(corrected_Stds_explant_panel_1, explant_panel_2,explant_panel_3)

#fixing the greek letters!!
explant_all_panels$Assay <-recode(explant_all_panels$Assay,
                                      "IFN-α2a" = paste("IFN-\U03B1","2a", sep = ""),
                                      "IL-1α" = "IL-1\U03B1",
                                      "MIP-1α"="MIP-1\U03B1",
                                      "TNF-α" = "TNF-\U03B1",
                                      "MIP-3α" = "MIP-3\U03B1",
                                      "MIP-1β" = "MIP-1\U03B2",
                                      "IL-1β" = "IL-1\U03B2",
                                      "IFN-γ" = "IFNg")

#Just keep the columns that I care about

explant_all_panels <- explant_all_panels %>%
  select(Assay,Sample, Sample.Group, Plate.Name, X..Recovery, Concentration, Calc..Concentration, Detection.Range, Fit.Statistic..RSquared,Spot)
```

## Checking Standard Recovery and Curves

Here are plots of the % recovery (extrapolated concentration/ known concentration x 100) for the standards that were either within or above the detection range, with lines at the "good recovery" limits of 80 and 120% recovery.

The R^2^ values for goodness-of-fit for the standard curves were all >0.99

```{r standards, echo=FALSE, message = FALSE, warning = FALSE}

explant_stds_all_panels<-filter(explant_all_panels, Sample.Group == "Standards")

#which standards are out of det. range?
standardsOutOfRange<-explant_stds_all_panels %>%
  filter(Detection.Range != "In Detection Range")%>%
  select(Plate.Name,Sample, Assay, Detection.Range, X..Recovery, Calc..Concentration)

#looking at which samples have STDs below the fit curve or below the detection range (i.e. you can't get numbers for % recovery)

explant_stds_all_panels_below<- explant_stds_all_panels%>%
  filter(Detection.Range == "Below Fit Curve Range" | Detection.Range == "Below Detection Range") %>%
  group_by(Sample)%>%
  summarise("Number of Samples Below Fit Curve Range/Detection Range" = n())

#The samples that actually give me #'s for percent recovery ( i.e. all but the "below fit curve range" and "below detection range" ones)

explant_stds_all_panels_percent_recov<- explant_stds_all_panels %>% 
  arrange(Sample,Assay)%>%
  filter(Detection.Range != "Below Fit Curve Range" & Detection.Range!= "Below Detection Range")%>%
  select(Plate.Name,Sample, Assay,Detection.Range, X..Recovery,Fit.Statistic..RSquared, Concentration, Calc..Concentration)

#plotting percent recovery

ggplot(explant_stds_all_panels_percent_recov,aes( x = Sample, y = X..Recovery)) +
  geom_point(aes(shape = Detection.Range ),
             size = 2, alpha = 0.5)+
  scale_shape_manual(values = c("Above Fit Curve Range" = 1, 
                                "In Detection Range" = 16, 
                                "Below Detection Range" = 0))+
  facet_wrap(~Assay)+
  theme(axis.text.x = element_text(angle=25))+
  scale_y_continuous(limits = c(0, 250))+
  geom_abline(slope = 0, intercept = c(80, 120))+
  labs(y = "Percent Recovery")+
  theme(axis.text.x = element_text(size = rel(0.8)))

#checking the R squared goodness-of-fit for the standard curves
rSquared<-explant_stds_all_panels_percent_recov%>%
  group_by(Assay)%>%
  summarize(R_Squared = mean(Fit.Statistic..RSquared))

```

## Caveats about Standards

* Standards with values *below the detection range* (<2.5x the standard deviation of the lowest standard) give unreliable concentrations, so I did not include those in the plots.

* When standards are *below the curve fit* (outside the range of the standard curve), the software does not calculate concentrations so you can't get a percent recovery.

* Most of the Standards that fell into either of these categories were from STD-07 or STD-08 except for two that were STD-06:

`r pander(explant_stds_all_panels_below)`

```{r break up sample name, echo=FALSE, message = FALSE, warning = FALSE}

#Making a df that contains just the non-standards

explant_all_panels_noStds <- explant_all_panels %>%
  filter(Sample.Group!="Standards")

# I want to be able to group the data by timepoint and donor and virus so I need to break up components of the the sample name and make new columns with just the Donor, Timepoint and Virus

#Split samples names on the "-"
donorTimeVirus<-colsplit(explant_all_panels_noStds$Sample,"-", c("Donor","Timepoint","Virus"))

#Add a column for Sample dilution (all samples were diluted 1:2)

donorTimeVirusDil <- donorTimeVirus %>%
  mutate(Sample_Dilution = "1:2")

#bind the donorTimeVirusDil to allPanels

explant_all_panels_noStds<-cbind(donorTimeVirusDil, explant_all_panels_noStds)

#Make things factors so they plot nicely
explant_all_panels_noStds$Donor <- as.factor(explant_all_panels_noStds$Donor)

explant_all_panels_noStds$Calc..Concentration <-as.numeric(explant_all_panels_noStds$Calc..Concentration)

explant_all_panels_noStds$Virus <- as.factor(explant_all_panels_noStds$Virus)

explant_all_panels_noStds$Virus<-factor(explant_all_panels_noStds$Virus, levels = c("M", "V1","V2"))

explant_all_panels_noStds$Timepoint<-as.factor(explant_all_panels_noStds$Timepoint)
```

```{r checking reps, echo=FALSE, message = FALSE, warning = FALSE}

#checking to make sure that only the samples that SHOULD have 1 rep, show up as having just 1 rep (sample 324-T2-V1 where I ran out of sample before I could do the 2nd rep)

replicateSummary <- explant_all_panels_noStds %>%
group_by(Plate.Name,Sample, Assay)%>%
summarize(reps = n())

not2<-filter(replicateSummary,reps!="2")%>%
  arrange(Sample, Assay)

```

## Explant Supernatents: Detection Range

```{r checking detection range, echo=FALSE, message = FALSE, warning = FALSE}

#What number of samples are in each category of detection?
explant_all_panels_noStds_NumInEach <- explant_all_panels_noStds%>%
  mutate(Panel = ifelse(str_detect(Plate.Name, "panel_1"),"Panel 1",ifelse(str_detect(Plate.Name,"panel_2"),"Panel 2","Panel 3")))%>%
 group_by(Panel,Assay, Detection.Range)%>%
  summarise(Number_of_Samples = n())

#Let's plot that

ggplot(explant_all_panels_noStds_NumInEach, aes(x = Assay, y=Number_of_Samples))+
         geom_point(aes(color = Detection.Range),size = 2)+
  theme(axis.text.x = element_text(angle=25, size = 7))+
  labs(y = "Number of Samples", x = "Analyte")+
  facet_wrap(~Panel, scales = "free")+
  ggtitle("Number of Samples in each Detection Range per Analyte")

```

```{r multiplying by dilution factor, echo=FALSE, message = FALSE, warning = FALSE}

#I'm going to make a column for estimated stock concentration by multiplying the avgConc by 2 (since I did a 1:2 dilution on the samples) so I can better compare with the HVE sup concentrations.

explant_all_panels_noStds <- explant_all_panels_noStds%>%
  mutate(Estimated_Stock_Conc =  Calc..Concentration * 2)

```

```{r exclude belows, echo=FALSE, message = FALSE, warning = FALSE}

#I am also excluding samples that are "Below Detection Range" because this means that the signal was "statistically indistinguishable from the background or top of the curve"" and ..."The concentrations calculated are inherently unreliable."

explant_all_panels_noStds_noBelow <- explant_all_panels_noStds %>%
  filter(Detection.Range != "Below Fit Curve Range" & Detection.Range != "Below Detection Range")
```

## All Samples Estimated Stock Concentrations

* I multiplied by the dilution factor so this is an estimated concentration of the stock samples.

* Note that this is on a log(10) scale; most concentrations are a lot lower than the high ones.

```{r  all sample concentrations, echo=FALSE, message = FALSE, warning = FALSE}

#Plot of all samples and their concentrations

ggplot(explant_all_panels_noStds_noBelow,aes( x = Timepoint, y = Estimated_Stock_Conc))+
         geom_point(aes(color = Assay),position = position_swarm (0.5),size = 2)+
  scale_y_log10()+
  labs(y = "Estimated Stock Concentration pg/ml")+
  facet_wrap(~Virus)

```

```{r averaging reps, echo=FALSE, message = FALSE, warning = FALSE}

#averaging the replicates

#some of the samples had one rep that was below detection, so I am excluding those so I will be able to use the remaining rep as the "average" for that sample.

avgConc_explant_all_panels_noStds_noBelow <- explant_all_panels_noStds_noBelow %>%
group_by(Sample, Assay, Donor, Timepoint,Virus)%>%
summarize(avgEstConc = mean(Estimated_Stock_Conc))%>%
  ungroup()

#Looking at the number of samples in each sample "set"(Same Donor and Timepoint). Some are missing V1 or V2. or both!! I don't want the ones with just one sample in the set.

numberInSets<-avgConc_explant_all_panels_noStds_noBelow %>%
         group_by(Donor,Timepoint,Assay)%>%
  select(Sample,Donor, Timepoint,Assay)%>%
  summarize(numInSet = n())

#now I'll merge the avgConc df with this numberInSets info created above so I can then filter out samples that have <2 samples in the set

avgConc_explant_all_panels_noStds_noBelow <- merge(avgConc_explant_all_panels_noStds_noBelow, numberInSets, by = c("Donor", "Timepoint", "Assay"))

#now exclude the samples where there not a "full set", i.e. 1 sample only in the set

avgConc_explant_all_panels_noStds_noBelow_fullSet <- avgConc_explant_all_panels_noStds_noBelow %>%
  filter(numInSet != 1)

```

## Plots of analyte concentrations

We did a 1:5 dilution of the samples so I multiplied the concentrations by 5 to get an estimate of the stock suparnatent concentration. Black points show the means.

## Caveats
Samples omitted from the plots include:

* Concentrations below the fit curve range. Values cannot be extrapolated when the data was *below* the fit curve range. However,the MSD software can extrapolate values that are *above* the fit curve range if the curve is linear at the top.

* Samples for which there was only data for a Mock sample and neither of the virus conditions.

Also note:
If a sample had one rep that was below detection and another that was in range, I used the remaining "in range" rep as the "average" for that sample.

```{r  individual analyte plots, echo=FALSE, message = FALSE, warning = FALSE}

ggplot(avgConc_explant_all_panels_noStds_noBelow_fullSet, aes(x = Timepoint, y = avgEstConc))+
   geom_point(aes(color = Donor),position = position_swarm(0.07), alpha = 0.75, size = 1.5)+
  scale_y_log10()+
   labs( y = "Estimated Stock Concentration pg/ml")+
   facet_grid(Assay~Virus, scales = "free_x")+
   theme(axis.text.x = element_text(size = 11))

################## another way of making the plots #####

# #make a list of the analytes
# 
# #I'm using a unique vector here instead of the levels of Assay because I don't want to include IFNa2a OR Il-7. There weren't any data points in the detection range for those analytes (so I don't want to try to make a plot of it)
# 
# 
# analyteList<-as.list(unique(avgConc_explant_all_panels_noStds_noBelow$Assay))
# 
# 
# 
# 
# #write a function to make a plot for each analyte
# 
# plotAnalytes <- function (Analyte){
#    avgConc_explant_all_panels_noStds_noBelow_fullSet %>%
#      filter(Assay == Analyte)%>%
#      
#      ggplot( aes(x = Timepoint, y = Estimated_Stock_Conc))+
#    
#    geom_point(aes(color = Donor ), alpha = 0.5, size = 3)+
#   geom_line(aes(group = Donor, color = Donor), alpha = 0.5)+
#     stat_summary(fun.y = mean, geom = "point")+
#     stat_summary(fun.y = mean, geom = "line",aes(group = 1))+
#    
#    scale_shape_manual(values = c("Above Fit Curve Range" = 1, "Below Detection Range" = 1, "In Detection Range" = 16))+
#   
#    scale_y_log10()+
#    
#    labs( y = "Concentration pg/ml")+
#    
#   facet_wrap(~Virus, scales = "free_x")+
#    
#    theme(axis.text.x = element_text(size = 11))+
#      ggtitle(Analyte)
# }
# 
# 
# 
# lapply(analyteList, FUN = plotAnalytes)

```