### 2016-09-22 MSD_panels_1_2_3_HVE_sup_analysis.Rmd
get logFC ratio by doing difference of log mock subtracted for 24hr and 3 hr (bcause log(a) -log(b) = log(a/b)

---
title: "MSD panels 1, 2 and 3 HVE Supernatant"
output: html_document
---

## Experiment overview

These data are from running MSD analysis of supernatents from selected ex vivo vaginal epithelial cells from three donors. The explants were infected with HSV2 strains 186, SD90 or Mock infected by Lamar Fleming and Gabriella Fenkart and supernatents were collected at 3 timepoints post-infection: 3hrs, 8hrs and 24hrs.

We used 3 different MSD panels containing a total of 20 analytes analyzed for 24 samples. Based on results from running test samples and from running the explant supernatent samples, we decided to dilute all the supernants 1:2 (plate 1) and 1:5 (plate 2) with MSD's diluent 43 before running the plates. All samples were run in duplicate.

```{r read in data, message=FALSE, warning = FALSE, echo = FALSE}

library(dplyr)
library(ggplot2)
library(pander)
library(reshape2)
library(stringr)
library(swarm)

#panel 1
# 128 wells x 9 analytes = 1152 obs
HVE_panel_1 <- read.csv("../raw_data/17Aug16_HVE_sup_panel_1_plate_1_and_2.csv", skip = 1)

#panel 2
#128 wells x 8 analytes = 1024 obs
HVE_panel_2 <- read.csv("../raw_data/17Aug16_HVE_sup_panel_2_plate_1_and_2.csv", skip = 1)

#panel 3
#128 wells x 3 analytes = 384 obs

HVE_panel_3 <- read.csv("../raw_data/18Aug16_HVE_sup_panel_3_plate_1_and_2.csv", skip = 1)

#combine data from all 3 panels

HVE_all_panels <-rbind(HVE_panel_1, HVE_panel_2, HVE_panel_3)

#Fix the greek

HVE_all_panels$Assay <-recode(HVE_all_panels$Assay,
                                      "IFN-α2a" = paste("IFN-\U03B1","2a", sep = ""),
                                      "IL-1α" = "IL-1\U03B1",
                                      "MIP-1α"="MIP-1\U03B1",
                                      "TNF-α" = "TNF-\U03B1",
                                      "MIP-3α" = "MIP-3\U03B1",
                                      "MIP-1β" = "MIP-1\U03B2",
                                      "IL-1β" = "IL-1\U03B2",
                                      "IFN-γ" = "IFNg",
                              "MIG/CXCL9" = "MIG")

#Just keep the columns that I care about

HVE_all_panels <- HVE_all_panels %>%
  select(Assay,Sample, Sample.Group, Plate.Name, X..Recovery, Concentration, Calc..Concentration, Detection.Range, Fit.Statistic..RSquared,Spot)

```

## Checking Standard Recovery and Curves

Here are plots of the % recovery (extrapolated concentration/ known concentration x 100) for the standards that were either within or above the detection range, with lines at the "good recovery" limits of 80 and 120% recovery.

The R^2^ values for goodness-of-fit for the standard curves were all >0.99

```{r standards,fig.height = 7, fig.width = 12, echo=FALSE, message = FALSE, warning = FALSE}

HVE_stds_all_panels<-filter(HVE_all_panels, Sample.Group == "Standards")

#which standards are out of det. range?
standardsOutOfRange<-HVE_stds_all_panels %>%
  filter(Detection.Range != "In Detection Range")%>%
  select(Plate.Name,Sample, Assay, Detection.Range, X..Recovery, Calc..Concentration)

#looking at which samples have STDs below the fit curve or below the detection range (i.e. you can't get numbers for % recovery)

HVE_stds_all_panels_below<- HVE_stds_all_panels%>%
  filter(Detection.Range == "Below Fit Curve Range" | Detection.Range == "Below Detection Range") %>%
  group_by(Sample)%>%
  summarise("Number of Samples Below Fit Curve Range/Detection Range" = n())

#The samples that actually give me #'s for percent recovery ( i.e. all but the "below fit curve range" and "below detection range" ones)

HVE_stds_all_panels_percent_recov<- HVE_stds_all_panels %>% 
  arrange(Sample,Assay)%>%
  filter(Detection.Range != "Below Fit Curve Range" & Detection.Range!= "Below Detection Range")%>%
  select(Plate.Name,Sample, Assay,Detection.Range, X..Recovery,Fit.Statistic..RSquared, Concentration, Calc..Concentration)

#plotting percent recovery

ggplot(HVE_stds_all_panels_percent_recov,aes( x = Sample, y = X..Recovery)) +
  geom_point(aes(shape = Detection.Range ),
             size = 2, alpha = 0.5)+
  scale_shape_manual(values = c("Above Fit Curve Range" = 1, 
                                "In Detection Range" = 16, 
                                "Below Detection Range" = 0))+
  facet_wrap(~Assay)+
  theme(axis.text.x = element_text(angle=25))+
  scale_y_continuous(limits = c(0, 265))+
  geom_abline(slope = 0, intercept = c(80, 120))+
  labs(y = "Percent Recovery")+
  theme(axis.text.x = element_text(size = 9))+
  ggtitle("Percent Recovery of Standards")

#checking the R squared goodness-of-fit for the standard curves
rSquared<-HVE_stds_all_panels_percent_recov%>%
  group_by(Assay)%>%
  summarize(R_Squared = mean(Fit.Statistic..RSquared))
```

## Caveats about Standards

* Standards with values *below the detection range* (<2.5x the standard deviation of the lowest standard) give unreliable concentrations, so I did not include those in the plots.

* When standards are *below the curve fit* (outside the range of the standard curve), the software does not calculate concentrations so you can't get a percent recovery.

* Most of the Standards that fell into either of these categories were from STD-07 or STD-08 except for five that were STD-06:

`r pander(HVE_stds_all_panels_below)`

```{r break up sample name, echo=FALSE, message = FALSE, warning = FALSE}
#Making a df that contains just the non-standards

HVE_all_panels_noStds <- HVE_all_panels %>%
  filter(Sample.Group!="Standards")

# I want to be able to group the data by timepoint and donor and virus so I need to break up components of the the sample name and make new columns with just the Donor, Timepoint, Virus and Sample_Dilution.

#Split samples names on the "-"
donorTimeVirusDil<-colsplit(HVE_all_panels_noStds$Sample,"-", c("Donor","Timepoint","Virus", "Sample_Dilution") )

#bind the split col back to allPanels

HVE_all_panels_noStds<-cbind(donorTimeVirusDil, HVE_all_panels_noStds)

#Make things factors so they plot nicely
HVE_all_panels_noStds$Donor <- as.factor(HVE_all_panels_noStds$Donor)

HVE_all_panels_noStds$Calc..Concentration <-as.numeric(HVE_all_panels_noStds$Calc..Concentration)

HVE_all_panels_noStds$Virus <- as.factor(HVE_all_panels_noStds$Virus)

HVE_all_panels_noStds$Virus<-factor(HVE_all_panels_noStds$Virus, levels = c("M", "V1","V2"))

HVE_all_panels_noStds$Timepoint<-as.factor(HVE_all_panels_noStds$Timepoint)

HVE_all_panels_noStds$Sample_Dilution<-as.factor(HVE_all_panels_noStds$Sample_Dilution)
```

## HVE Supernatents: Detection Range

The following plots show that the 1:2 dilution of the samples resulted in more samples being in "In Detection Range."

```{r checking detection range, fig.height=8, fig.width =12, echo=FALSE, message = FALSE, warning = FALSE}

#What number of samples are in each category of detection? Also adding a column to differentiate which analyte is in which panel, mostly for facetting reasons.

HVE_all_panels_noStds_NumInEach <- HVE_all_panels_noStds%>%
  mutate(Panel = ifelse(str_detect(Plate.Name, "panel_1"),"Panel 1",ifelse(str_detect(Plate.Name,"panel_2"),"Panel 2","Panel 3")))%>%
 group_by(Plate.Name,Sample_Dilution,Panel,Assay, Detection.Range)%>%
  summarise(Number_of_Samples = n())

#Let's plot that

ggplot(HVE_all_panels_noStds_NumInEach, aes(x = Assay, y=Number_of_Samples))+
         geom_point(aes(color = Detection.Range),size = 2)+
  theme(axis.text.x = element_text(angle=25, size = 11))+
  labs(y = "Number of Samples", x = "Analyte")+
  facet_wrap(Sample_Dilution~Panel, scales = "free")+
  ggtitle("Number of Samples in each Detection Range per Analyte")
```

```{r multiplying by dilution factor, echo=FALSE, message = FALSE, warning = FALSE}

#I'm going to make a column for estimated stock concentration by multiplying the avgConc by 2 or 5 (since I did a 1:2 and 1:5 dilution on the samples) so I can better compare with the explant sup concentrations.

HVE_all_panels_noStds <- HVE_all_panels_noStds%>%
  mutate(Estimated_Stock_Conc = ifelse(
    Sample_Dilution == "1:2", Calc..Concentration * 2,
    ifelse(Sample_Dilution == "1:5", Calc..Concentration *5, NA)
  ))

```

```{r adjust belows, echo=FALSE, message = FALSE, warning = FALSE}
#Samples that are "Below Detection Range" means that the signal was "statistically indistinguishable from the background or top of the curve"" and ..."The concentrations calculated are inherently unreliable."

#HOWEVER, we still want to show the fact that those samples had really low concentrations. So, I am going to set all those samples to 0.001 so they show up on plots.

HVE_all_panels_noStds_AdjBelow <- HVE_all_panels_noStds %>%
  mutate(Estimated_Stock_Conc = ifelse(str_detect(Detection.Range,"Below"),0.01, Estimated_Stock_Conc))
```

## Estimated Sample Concentration Distribution

* I multiplied by the dilution factors so this is an estimated concentration of the stock samples.

* Note that this is on a log(10) scale; most concentrations are a lot lower than the high ones.

```{r filter for 1:2 MOSTLY, echo=FALSE, message = FALSE, warning = FALSE}
### NOTE ### I'm going to filter for just the 1:2 dilution samples here since they gave more in-detection. its ca''ed short because theres only 1:2 data

HVE_all_panels_noStds_AdjBelow_short<-HVE_all_panels_noStds_AdjBelow %>%
  filter(Sample_Dilution =="1:2")

#There were a few samples where the 1:2 dilution was above the fit curve range but it was in range for 1:5. I am going to extract those samples and add them to the df I made above, along with their corresponding Mock samples (which are also in range

IL1RA_4_T3<-HVE_all_panels_noStds_AdjBelow%>%
  filter(Assay == "IL1-RA", Donor == 4, Timepoint == "T3", Sample_Dilution=="1:5")

IL1RA_2_T2<-HVE_all_panels_noStds_AdjBelow%>%
  filter(Assay == "IL1-RA", Donor == 2, Timepoint == "T2", Sample_Dilution=="1:5")

HVE_all_panels_noStds_AdjBelow_short <- HVE_all_panels_noStds_AdjBelow_short %>%
  filter(!(Assay == "IL1-RA" & Donor == 2 & 
             Timepoint == "T2" & Sample_Dilution=="1:2")) %>% 
  filter(!(Assay == "IL1-RA" & Donor == 4 & 
             Timepoint == "T3" & Sample_Dilution=="1:2"))

HVE_all_panels_noStds_AdjBelow_short <- rbind(
  HVE_all_panels_noStds_AdjBelow_short, IL1RA_4_T3, IL1RA_2_T2)

```

```{r make adjustments for nice plotting, echo=FALSE, message = FALSE, warning = FALSE}

HVE_all_panels_noStds_AdjBelow_short<-HVE_all_panels_noStds_AdjBelow_short %>%
  mutate(Timepoint = ifelse(Timepoint == "T1", "3",
                            ifelse(Timepoint=="T2","8","24")))

#factor to get order right
HVE_all_panels_noStds_AdjBelow_short$Timepoint<-factor(HVE_all_panels_noStds_AdjBelow_short$Timepoint,levels = c("3","8","24"))

HVE_all_panels_noStds_AdjBelow_short<-HVE_all_panels_noStds_AdjBelow_short%>%
  mutate(Virus = ifelse(Virus == "V1", "SD90",
                        ifelse(Virus == "V2", "186","Mock")))

HVE_all_panels_noStds_AdjBelow_short$Virus<-factor(HVE_all_panels_noStds_AdjBelow_short$Virus,levels = c("Mock","SD90","186"))

```

```{r  all sample concentrations,fig.height = 6, fig.with = 10,  echo=FALSE, message = FALSE, warning = FALSE}

#Plot of all samples and their concentrations

ggplot(HVE_all_panels_noStds_AdjBelow_short,aes( x = Timepoint, y = Estimated_Stock_Conc))+
         geom_point(aes(color = Assay),position = position_swarm (0.5),size = 2)+
  scale_y_log10()+
  labs(y = "Estimated Stock Concentration pg/ml")+
  facet_wrap(~Virus)+
  ggtitle("Estimated Sample Concentration Distribution by Treatment and Timepoint")

```

```{r avg reps, echo=FALSE, message = FALSE, warning = FALSE}

#First change the "Below..." entries to just say "Below", since I don't care which "Below..." it was for my purposes.

HVE_all_panels_noStds_AdjBelow_short <- HVE_all_panels_noStds_AdjBelow_short %>%
  mutate(Detection.Range = ifelse(str_detect(Detection.Range, "Below"), "Below", as.character(Detection.Range)))

#which samples have discordant replicates in terms of det.range?
repsInRange<-HVE_all_panels_noStds_AdjBelow_short%>%
  group_by(Sample, Assay, Detection.Range)%>%
  summarize(Number_Reps = n())

#merge the repsInRange data with the HVE_all_panels_noStds_1to2_AdjBelow data by the Sample and Assay

#Result is a df where the samples with discordant replicates in terms of Detection.Range. The samples with "1" in the Number_of_Reps colimn are the ones where I want to use the data from the rep that was in/above range for the avg, NOT the average between the in-range rep and the below-range rep. Recall that I assigned all below range/curve reps a value of 0.01. If both reps were below, ok to use calculate the average of them.

# select fewer cols for readability.

mergeRepsInRange <- HVE_all_panels_noStds_AdjBelow_short %>%
  merge(.,repsInRange, by = c("Sample", "Assay", "Detection.Range"))%>%
  select(Sample,Timepoint,Donor,Virus,Assay,Detection.Range, Estimated_Stock_Conc, Number_Reps, Plate.Name, Sample_Dilution)

#The  rows of data that I want will fit the following criteria:

#Detection.Range = Below and Number_reps = 2 (both reps are below)
#Detection.Range != Below and Number_reps = 2 (both reps are good)
#Detection.Range != Below and Number_reps = 1 (The in/above range rep from a discordant set)

#I am doing this by filtering for the things I want to keep, then combining those dfs, rather than writing one expression to say what to exclude. (mostly because I couldnt figure it out...)

bothBad <- mergeRepsInRange %>%
  filter(Detection.Range =="Below"& Number_Reps ==2)

discordGood<- mergeRepsInRange %>%
  filter(Detection.Range == "In Detection Range" & Number_Reps ==1)

bothGood <- mergeRepsInRange %>%
  filter(Detection.Range == "In Detection Range"& Number_Reps == 2)

#combine all the reps, including the data the I extracted from the 1:5 dilution for IL1-RA

OKtoAvg <- rbind(bothBad,discordGood, bothGood)

#averaging the replicates. Should have 944 reps - 26 single reps = 918 reps

#918 reps /2 =459 averages + 26 averages from single reps = 485 values
avgConc <- OKtoAvg %>%
group_by(Sample, Assay, Donor, Timepoint,Virus, Sample_Dilution)%>%
summarize(avgEstConc = mean(Estimated_Stock_Conc))%>%
  ungroup()

#Looking at the number of samples in each sample "set"(Same Donor and Timepoint). Some are missing V1 or V2. or both!! I don't want the ones with just one sample in the set.

numberInSets<-avgConc %>%
         group_by(Donor,Timepoint,Assay)%>%
  select(Sample,Donor, Timepoint,Assay)%>%
  summarize(numInSet = n())

#all complete sets!

```

```{r merge and subtract mock, fig.width = 8, fig.height= 13, echo=FALSE}

#I want to subtract out the mock data from the corresponding v1 and V2 data.

#extract out Mock samples

mock <- avgConc %>%
  filter(Virus == "Mock")

# remove mock from the rest of the data

no_mock_avgConc <-
  avgConc %>%
  filter(Virus !="Mock")

mergeMock<-merge(no_mock_avgConc, mock, by = c("Donor","Timepoint","Assay", "Sample_Dilution"), all.x = TRUE, all.y = TRUE)

#check for NA's (places where there was a mock sample but no corresponding V1 or V2)
whichNAs<-mergeMock[which(is.na(mergeMock$Sample.x)),]
#no NA's!

#Fix up the colnames:

names(mergeMock)[c(5:7,10)]<- c("Sample","Virus","avgEstConcVirus","avgEstConcMock")

#do the subtraction, with the result being the difference between the log10 of virus - log 10 Mock
mergeMock <- mergeMock %>%
  mutate(logMockSubtracted = log10(avgEstConcVirus) - log10(avgEstConcMock))

```

Samples are multiplied by their dilution factor to get estimated concentrations.

If a sample had one rep that was below detection and another that was in range, I used the remaining "in range" rep as the "average" for that sample.

```{r setup for individual analyte HEATMAP, echo=FALSE}

#I want to make a heatmap that has the following data

#NO 8hr timepoint data

#tiles represent foldchange of Mock subtracted 24hr data over mock-subtracted 3hr. So if 24hr>3hr, you get a ratio <1 and if 24hr <3hr you get a ratio >1

#Want gradient intensity to max out at FC of 10.

#Exclude 8hr data.

no8mergeMock <- mergeMock %>%
  filter(Timepoint != "8")

# Extract the 3hr data and select a few columns

timepoint3<- no8mergeMock %>%
  filter(Timepoint == "3")%>%
  select(Donor, Assay, Sample, Virus, logMockSubtracted)

#extract 24 hr data and select a few columns
timepoint24<- no8mergeMock %>%
  filter(Timepoint =="24")%>%
  select(Donor, Assay, Sample, Virus, logMockSubtracted)

```

```{r merge 24 and 3, check NAs, echo=FALSE}
#merge 24hr and 3hr data

merge24And3<- merge(timepoint24,timepoint3, by  = c("Donor","Assay","Virus"), all.x = TRUE, all.y = TRUE)

#There are NO rows where we have data for 24hr but NOT for 3hr.
which24no3<-merge24And3[which(is.na(merge24And3$Sample.y)),]

#there are NO rows where we have data for 3hr but not for 24hr
which3no24<-merge24And3[which(is.na(merge24And3$Sample.x)),]

#Fix up colnames      
names(merge24And3)[4:7]<- c("Sample.24", "logMockSubtracted.24","Sample.3", "logMockSubtracted.3")
```

```{r get ratio, check NAs, echo=FALSE}

#Do the subtraction to get a foldchange.

merge24And3 <- merge24And3 %>%
  mutate(logMockSub24_minus_logMockSub3 = logMockSubtracted.24-logMockSubtracted.3)
```

```{r setup logFC heatmap, check NAs, echo=FALSE}

#I want the scale to have max intensity a 10-fold change in either direction (log10(10)=1 on my scale and log10(0.1)= -1 on my scale) so I am going to set all values >=1 as 1 and all values <=-1 as -1.

#set all values >=1 as 1

merge24And3 <- merge24And3 %>%
  mutate(max10fold =
           ifelse(
             logMockSub24_minus_logMockSub3 >=1,
         1,
           ifelse(
             logMockSub24_minus_logMockSub3 <=-1,
         -1,logMockSub24_minus_logMockSub3)
           )
  )

```

```{r logFC heatmap, check NAs, echo=FALSE}

ggplot(merge24And3, aes(x = Donor, y = Assay))+
  geom_tile(aes(fill = max10fold), color = "gray")+
  scale_fill_gradient2(low = "green", mid = "white", high = "red", midpoint = 0)+
  facet_wrap(~Virus)+
  ggtitle("log fold change of concentrations at 24hrs over 3 hrs")
```

```{r trying pheatmap, check NAs, echo=FALSE}

#trying with pheatmap
# library(pheatmap)
# library(tidyr)
# 
# #trying with pheatmap since ggplot is funky looking and I want zero in the middle
# 
# #pheatmap need it as numeric matrix and can't facet so I'll do v186 separately from sd90
# 
# SD90heatmap<-merge24And3 %>%
#   filter(Virus == "SD90")%>%
#   select(Donor, Assay, max10fold)
# 
# 
# matrixSD90heatmapSpread<-SD90heatmap%>%
#   spread(key = Donor, value = max10fold)%>%
#   select(-Assay)%>%
#   as.matrix()
# 
# 
# rownames(matrixSD90heatmapSpread) <-SD90heatmapSpread$Assay
# 
# 
# pheatmap(matrixSD90heatmapSpread)
#

test = matrix(rnorm(200), 20, 10)
test[1:10, seq(1, 10, 2)] = test[1:10, seq(1, 10, 2)] + 3
test[11:20, seq(2, 10, 2)] = test[11:20, seq(2, 10, 2)] + 2
test[15:20, seq(2, 10, 2)] = test[15:20, seq(2, 10, 2)] + 4
colnames(test) = paste("Test", 1:10, sep = "")
rownames(test) = paste("Gene", 1:20, sep = "")

pheatmap(test)

```