### 2016-09-23 MSD_panels_1_2_3_explant_sup_analysis.Rmd
added scatter plots back to HVE file and edited plot titles in both files

---
title: "MSD panels 1, 2 and 3 Explant Supernatant"
output: github_document
---

## Experiment overview

These data are from running MSD analysis of supernatents from selected vaginal explant samples from seven donors. The explants were infected with HSV2 strains 186, SD90 or Mock infected by Lamar Fleming and Gabriella Fenkart and supernatents were collected at 3 timepoints post infection: 3hrs, 8hrs and 24hrs.

We used 3 different MSD panels containing a total of 20 analytes analyzed for 46 samples (see caveat). Based on results from running test samples, we decided to dilute all the supernants 1:5 with MSD's diluent 43 before running the plates. All samples (except 1, see caveats) were run in duplicate.

## Caveats

Due to limited sample volume, we had to make the following omissions:

Panel 2: Only 1 rep for sample 324-T2-V1

Panel 3: Did not run samples 324-T2-V1 or 324-T2-V2.

Due to a possible pipetting error, we were not able to use the data from the standards that we ran for panel 1. Instead, we used the standards from a different run of the same panel to calculate concentrations for the experimental samples (see explant_sup_README for details).

```{r data read-in, echo=FALSE, message = FALSE, warning = FALSE}

library(dplyr)
library(ggplot2)
library(pander)
library(reshape2)
library(stringr)
library(swarm)

#reading in data from the panel 1

#remember that this data table contains data from the 17Aug16_HVE_sup_panel_1_plate_1. I am only going to use the Standards data from that plate (not the data) and not the standards from either of the plates for panel 1 run 10Aug16 since we had what appears to be a pipetting error (2nd replicate is more concentrated for all wells).

explant_panel_1_and_HVE_panel_1_plate_1 <- read.csv("../raw_data/10Aug16_explant_sup_panel_1_plate_1_and_2.csv", skip = 1)

#Filter to ONLY keep the standards from the plate called "HVE_sup_panel_1_plate_1"

StdsToUse_explant_panel_1 <- explant_panel_1_and_HVE_panel_1_plate_1 %>%
  filter(Sample.Group == "Standards")%>%
  filter(Plate.Name == "HVE_sup_panel_1_plate_1")

#Since I told Discovery Workbench software that I wanted to use the standards from the 17Aug16_HVE_sup_panel_1_plate_1 as "Global Standards" (i.e. use them to calculate concentrations for all plates in the experiment)it seems to have provided me with 3 copies of those samples in the data table (1 for each plate I guess). I only need data from the unique entries, so I will filter for just unique rows in this df to exclude copies.

# should have 8 stds x 9 analytes x 2 reps = 144 obs
StdsToUse_explant_panel_1<-unique(StdsToUse_explant_panel_1)

#Sanity check: there should only be 2 entries for each unique instance of Sample (STD-0x) and assay. 
checkStdReps<-StdsToUse_explant_panel_1 %>%
         group_by(Sample,Assay)%>%
         summarise(n())

#Remove the 17Aug16_HVE_sup_panel_1_plate_1 non-STD data from the df and leave ONLY explant panel 1 data. Remaining should be:
#94 wells from plate 1 + 30 wells from plate 2 x 9 analytes = 1116 observations

explant_panel_1 <-explant_panel_1_and_HVE_panel_1_plate_1 %>%
  filter(Plate.Name != "HVE_sup_panel_1_plate_1")

#now remove the unwanted Std samples from the explant data. There should be 32 Std wells x 9 analytes = 288
# 1116 total - 288 = 828 observations

No_Stds_explant_panel_1 <-explant_panel_1 %>%
  filter(Sample.Group != "Standards")

#now add in the unique Stds from 17Aug16_HVE_sup_panel_1_plate_1

corrected_Stds_explant_panel_1 <- rbind(No_Stds_explant_panel_1,StdsToUse_explant_panel_1)

# Data from panels 2 and 3

#panel 2 should have 123 wells (left out one rep) x 8 analytes = 984
explant_panel_2<-read.csv("../raw_data/10Aug16_explant_sup_panel_2_plate_1_and_2.csv", skip = 1)

#panel 3: 120 wells (left out 2 samples, 4 reps) x 8 analytes =360

explant_panel_3<-read.csv("../raw_data/11Aug16_explant_sup_panel_3_plate_1_and_2.csv", skip = 1)

#Combining the data from all 3 panels: 972 +984 + 360= 2316

explant_all_panels<- rbind(corrected_Stds_explant_panel_1, explant_panel_2,explant_panel_3)

#fixing the greek letters!!
explant_all_panels$Assay <-recode(explant_all_panels$Assay,
                                      "IFN-α2a" = paste("IFN-\U03B1","2a", sep = ""),
                                      "IL-1α" = "IL-1\U03B1",
                                      "MIP-1α"="MIP-1\U03B1",
                                      "TNF-α" = "TNF-\U03B1",
                                      "MIP-3α" = "MIP-3\U03B1",
                                      "MIP-1β" = "MIP-1\U03B2",
                                      "IL-1β" = "IL-1\U03B2",
                                      "IFN-γ" = "IFNg",
                                  "MIG/CXCL9" = "MIG")

#Just keep the columns that I care about

explant_all_panels <- explant_all_panels %>%
  select(Assay,Sample, Sample.Group, Plate.Name, X..Recovery, Concentration, Calc..Concentration, Detection.Range, Fit.Statistic..RSquared,Spot)
```

## Checking Standard Recovery and Curves

Here are plots of the % recovery (extrapolated concentration/ known concentration x 100) for the standards that were either within or above the detection range, with lines at the "good recovery" limits of 80 and 120% recovery.

The R^2^ values for goodness-of-fit for the standard curves were all >0.99

```{r standards,fig.height = 7, fig.width = 12, echo=FALSE, message = FALSE, warning = FALSE}

explant_stds_all_panels<-filter(explant_all_panels, Sample.Group == "Standards")

#which standards are out of det. range?
standardsOutOfRange<-explant_stds_all_panels %>%
  filter(Detection.Range != "In Detection Range")%>%
  select(Plate.Name,Sample, Assay, Detection.Range, X..Recovery, Calc..Concentration)

#looking at which samples have STDs below the fit curve or below the detection range (i.e. you can't get numbers for % recovery)

explant_stds_all_panels_below<- explant_stds_all_panels%>%
  filter(Detection.Range == "Below Fit Curve Range" | Detection.Range == "Below Detection Range") %>%
  group_by(Sample)%>%
  summarise("Number of Samples Below Fit Curve Range/Detection Range" = n())

#The samples that actually give me #'s for percent recovery ( i.e. all but the "below fit curve range" and "below detection range" ones)

explant_stds_all_panels_percent_recov<- explant_stds_all_panels %>% 
  arrange(Sample,Assay)%>%
  filter(Detection.Range != "Below Fit Curve Range" & Detection.Range!= "Below Detection Range")%>%
  select(Plate.Name,Sample, Assay,Detection.Range, X..Recovery,Fit.Statistic..RSquared, Concentration, Calc..Concentration)

#plotting percent recovery

ggplot(explant_stds_all_panels_percent_recov,aes( x = Sample, y = X..Recovery)) +
  geom_point(aes(shape = Detection.Range ),
             size = 2, alpha = 0.5)+
  scale_shape_manual(values = c("Above Fit Curve Range" = 1, 
                                "In Detection Range" = 16, 
                                "Below Detection Range" = 0))+
  facet_wrap(~Assay)+
  theme(axis.text.x = element_text(angle=25))+
  scale_y_continuous(limits = c(0, 250))+
  geom_abline(slope = 0, intercept = c(80, 120))+
  labs(y = "Percent Recovery")+
  theme(axis.text.x = element_text(size = 9))+
  ggtitle("Percent Recovery of Standards")

#checking the R squared goodness-of-fit for the standard curves
rSquared<-explant_stds_all_panels_percent_recov%>%
  group_by(Assay)%>%
  summarize(R_Squared = mean(Fit.Statistic..RSquared))

```

## Caveats about Standards

* Standards with values *below the detection range* (<2.5x the standard deviation of the lowest standard) give unreliable concentrations, so I did not include those in the plots.

* When standards are *below the curve fit* (outside the range of the standard curve), the software does not calculate concentrations so you can't get a percent recovery.

* Most of the Standards that fell into either of these categories were from STD-07 or STD-08 except for two that were STD-06:

`r pander(explant_stds_all_panels_below)`

```{r break up sample name, echo=FALSE, message = FALSE, warning = FALSE}

#Making a df that contains just the non-standards

explant_all_panels_noStds <- explant_all_panels %>%
  filter(Sample.Group!="Standards")

# I want to be able to group the data by timepoint and donor and virus so I need to break up components of the the sample name and make new columns with just the Donor, Timepoint and Virus

#Split samples names on the "-"
donorTimeVirus<-colsplit(explant_all_panels_noStds$Sample,"-", c("Donor","Timepoint","Virus"))

#Add a column for Sample dilution (all samples were diluted 1:5)

donorTimeVirusDil <- donorTimeVirus %>%
  mutate(Sample_Dilution = "1:5")

#bind the donorTimeVirusDil to allPanels

explant_all_panels_noStds<-cbind(donorTimeVirusDil, explant_all_panels_noStds)

#Make things factors so they plot nicely
explant_all_panels_noStds$Donor <- as.factor(explant_all_panels_noStds$Donor)

explant_all_panels_noStds$Calc..Concentration <-as.numeric(explant_all_panels_noStds$Calc..Concentration)

explant_all_panels_noStds$Virus <- as.factor(explant_all_panels_noStds$Virus)

explant_all_panels_noStds$Virus<-factor(explant_all_panels_noStds$Virus, levels = c("M", "V1","V2"))

explant_all_panels_noStds$Timepoint<-as.factor(explant_all_panels_noStds$Timepoint)
```

```{r checking reps, echo=FALSE, message = FALSE, warning = FALSE}

#checking to make sure that only the samples that SHOULD have 1 rep, show up as having just 1 rep (sample 324-T2-V1 where I ran out of sample before I could do the 2nd rep for panel 1. there wasnt enough of this sample to use any for panel 2 :( )

replicateSummary <- explant_all_panels_noStds %>%
group_by(Plate.Name,Sample, Assay)%>%
summarize(reps = n())

not2<-filter(replicateSummary,reps!="2")%>%
  arrange(Sample, Assay)

```

## Explant Supernatents: Detection Range

```{r checking detection range,fig.height = 4, fig.width = 12, echo=FALSE, message = FALSE, warning = FALSE}

#What number of samples are in each category of detection?
explant_all_panels_noStds_NumInEach <- explant_all_panels_noStds%>%
  mutate(Panel = ifelse(str_detect(Plate.Name, "panel_1"),"Panel 1",ifelse(str_detect(Plate.Name,"panel_2"),"Panel 2","Panel 3")))%>%
 group_by(Panel,Assay, Detection.Range)%>%
  summarise(Number_of_Samples = n())

#Let's plot that

ggplot(explant_all_panels_noStds_NumInEach, aes(x = Assay, y=Number_of_Samples))+
         geom_point(aes(color = Detection.Range),size = 2)+
  theme(axis.text.x = element_text(angle=25, size = 8))+
  labs(y = "Number of Samples", x = "Analyte")+
  facet_wrap(~Panel, scales = "free")+
  ggtitle("Number of Samples in each Detection Range per Analyte")

#Were there any samples where none of the assays seemed to have worked 
#i.e. All 40 were below?
explant_all_panels_noStds_NumInEach_bySample <- explant_all_panels_noStds%>%
 group_by(Sample, Detection.Range)%>%
  summarise(Number_of_Reps = n())%>%
  filter(str_detect(Detection.Range, "Below"))%>%
  group_by(Sample)%>%
  summarise(Number_of_Belows = sum(Number_of_Reps))
#None of these numbers are 40, so it is ok.

```

```{r multiplying by 5, echo=FALSE, message = FALSE, warning = FALSE}

#I'm going to make a column for estimated stock concentration by multiplying the avgConc by 5 (since I did a 1:5 dilution on the samples) so I can better compare with the HVE sup concentrations.

explant_all_panels_noStds <- explant_all_panels_noStds%>%
  mutate(Estimated_Stock_Conc =  Calc..Concentration * 5)

```

```{r adjust belows, echo=FALSE, message = FALSE, warning = FALSE}

#Samples that are "Below Detection Range" means that the signal was "statistically indistinguishable from the background or top of the curve" and ..."The concentrations calculated are inherently unreliable."

#HOWEVER, we still want to show the fact that those samples had really low concentrations. So, I am going to set all those samples to 0.001 so they show up on plots.

#which ones were those again?

explant_belows <- explant_all_panels_noStds %>%
  filter(str_detect(Detection.Range, "Below"))%>%
  group_by(Donor, Timepoint, Assay)%>%
  summarize(Number_of_samples = n())

#I'm going to set all of the "belows" to 0.001

explant_all_panels_noStds_AdjBelow <- explant_all_panels_noStds %>%
mutate(Estimated_Stock_Conc = ifelse(str_detect(Detection.Range,"Below"),0.001, Estimated_Stock_Conc))

```

##Estimated Sample Concentration Distribution

* I multiplied by the dilution factor so this is an estimated concentration of the stock samples.

* Note that this is on a log(10) scale; most concentrations are a lot lower than the high ones.

```{r make adjustments for nice plotting, echo=FALSE, message = FALSE, warning = FALSE}

explant_all_panels_noStds_AdjBelow<-explant_all_panels_noStds_AdjBelow %>%
  mutate(Timepoint = ifelse(Timepoint == "T1", "3",
                            ifelse(Timepoint=="T2","8","24")))

#factor to get order right
explant_all_panels_noStds_AdjBelow$Timepoint<-factor(explant_all_panels_noStds_AdjBelow$Timepoint,levels = c("3","8","24"))

explant_all_panels_noStds_AdjBelow<-explant_all_panels_noStds_AdjBelow%>%
  mutate(Virus = ifelse(Virus == "V1", "SD90",
                        ifelse(Virus == "V2", "186","Mock")))

explant_all_panels_noStds_AdjBelow$Virus<-factor(explant_all_panels_noStds_AdjBelow$Virus,levels = c("Mock","SD90","186"))

```

```{r  all sample concentrations,fig.height = 6, fig.with = 10, echo=FALSE, message = FALSE, warning = FALSE, results = 'asis'}

#Plot of all samples and their concentrations
#Note that this includes the samples that manually set to 0.001

ggplot(explant_all_panels_noStds_AdjBelow,aes( x = Timepoint, y = Estimated_Stock_Conc))+
         geom_point(aes(color = Assay),position = position_swarm (0.5),size = 2)+
  scale_y_log10()+
  labs(y = "Estimated Stock Concentration pg/ml")+
  facet_wrap(~Virus)+
  ggtitle("Estimated Sample Concentration Distribution by Treatment and Timepoint")

#("\n\n\\pagebreak\n")

```

```{r averaging reps, echo=FALSE, message = FALSE, warning = FALSE}

#First change the "Below..." entries to just say "Below", since I don't care which "Below..." it was for my purposes.

explant_all_panels_noStds_AdjBelow <- explant_all_panels_noStds_AdjBelow %>%
  mutate(Detection.Range = ifelse(str_detect(Detection.Range, "Below"), "Below", as.character(Detection.Range)))

#which samples have discordant replicates in terms of det.range?
repsInRange<-explant_all_panels_noStds_AdjBelow%>%
  group_by(Sample, Assay, Detection.Range)%>%
  summarize(Number_Reps = n())

#merge the repsInRange data with the explant_all_panels_noStds_AdjBelow data by the Sample and Assay

#Result is a df where the samples with dischordant replicates in terms of Detection.Range. The samples with "1" in the Number_of_Reps colimn are the ones where I want to use the data from the rep that was in/above range for the avg, NOT the average between the in-range rep and the below-range rep. Recall that I assigned all below range/curve reps a value of 0.01. If both reps were below, ok to use calculate the average of them.

# select fewer cols for readability.

mergeRepsInRange <- explant_all_panels_noStds_AdjBelow %>%
  merge(.,repsInRange, by = c("Sample", "Assay", "Detection.Range"))%>%
  select(Sample,Timepoint,Donor,Virus,Assay,Detection.Range, Estimated_Stock_Conc, Number_Reps, Plate.Name)

#HOWEVER: I know that for sample 324-T2-V1, I only had one replicate for the assays on panel 2, so for just this sample, I don't want to eliminate a rep that meets this exclusion criteria (IL-12p70) because I don't actually have a good rep to use as the average. I'll add it back in after filtering.

T2V1324<-mergeRepsInRange %>%
  filter(Sample == "324-T2-V1", Assay =="IL-12p70")

#The  rows of data that I want will fit the following criteria:

#Detection.Range = Below and Number_reps = 2 (both reps are below)
#Detection.Range != Below and Number_reps = 2 (both reps are good)
#Detection.Range != Below and Number_reps = 1 (The in/above range rep from a discordant set)

#I am doing this by filtering for the things I want to keep, then combining those dfs, rather than writing one expression to say what to exclude. (mostly because I couldnt figure it out...)

bothBad <- mergeRepsInRange %>%
  filter(Detection.Range =="Below"& Number_Reps ==2)

discordGood<- mergeRepsInRange %>%
  filter(Detection.Range != "Below"& Number_Reps ==1)

bothGood <- mergeRepsInRange %>%
  filter(Detection.Range != "Below"& Number_Reps == 2)

#combine all the reps

OKtoAvg <- rbind(bothBad,discordGood, bothGood,T2V1324)

#averaging the replicates
avgConc_explant_all_panels_noStds_AdjBelow <- OKtoAvg %>%
group_by(Sample, Assay, Donor, Timepoint,Virus)%>%
summarize(avgEstConc = mean(Estimated_Stock_Conc))%>%
  ungroup()

```

```{r  samples in sets, fig.width = 8, fig.height= 13, echo=FALSE}

#Looking at the number of samples in each sample "set"(Same Donor and Timepoint).

numberInSets<-avgConc_explant_all_panels_noStds_AdjBelow %>%
         group_by(Donor,Timepoint,Assay)%>%
  select(Sample,Donor, Timepoint,Assay)%>%
  summarize(numInSet = n())

#now I'll merge the avgConc df with this numberInSets info so I know which conditions are in each set.
avgConc_explant_all_panels_noStds_AdjBelow <- merge(avgConc_explant_all_panels_noStds_AdjBelow, numberInSets, by = c("Donor", "Timepoint", "Assay"))

# For 324-T2, there are ONLY mock samples
# There are no samples with just V1 and/or V2 and no mock

```

```{r  merge and subtract mock, fig.width = 8, fig.height= 13, echo=FALSE}

#I want to subtract out the mock data from the corresponding v1 and V2 data.

#extract out Mock samples

mock <- avgConc_explant_all_panels_noStds_AdjBelow %>%
  filter(Virus == "Mock")

# remove mock from the rest of the data

no_mock_avgConc_explant_all_panels_noStds_AdjBelow <-
  avgConc_explant_all_panels_noStds_AdjBelow %>%
  filter(Virus !="Mock")

mergeMock<-merge(no_mock_avgConc_explant_all_panels_noStds_AdjBelow, mock, by = c("Donor", "Assay", "Timepoint"), all.x = TRUE, all.y = TRUE)

#check for NA's (places where there was a mock sample but no corresponding V1 or V2)
whichNAs<-mergeMock[which(is.na(mergeMock$Sample.x)),]
#This is just donor 324 Il1ra, mig and rantes, which we already know has just mock.
# Remove the NAs for tidyness

mergeMock<-mergeMock%>%
  na.omit()%>%
  select(Donor, Assay, Timepoint, Sample.x, Virus.x, avgEstConc.x, numInSet.x, avgEstConc.y)

#Fix up the colnames

names(mergeMock)[4:8]<- c("Sample","Virus","avgEstConcVirus","numInSet","avgEstConcMock")

#do the subtraction, with the result being the difference between the log10 of virus - log 10 Mock
mergeMock <- mergeMock %>%
  mutate(logMockSubtracted = log10(avgEstConcVirus) - log10(avgEstConcMock))

```

We did a 1:5 dilution of the samples so I multiplied the concentrations by 5 to get an estimate of the stock supernatent concentration.

If a sample had one rep that was below detection and another that was in range, I used the remaining "in range" rep as the "average" for that sample.

```{r scatter plots of mocksubtracted, fig.width = 6,fig.height= 13, echo=FALSE}

ggplot(mergeMock, aes(x = Timepoint, y = logMockSubtracted))+
   geom_point(aes(color = Donor),position = position_swarm(0.1), alpha = 0.75, size = 3)+
   labs( y = "log 10 Estimated Stock Concentration pg/ml")+
   facet_grid(Assay~Virus, scales = "free_x")+
   theme(axis.text.x = element_text(size = 11))+
  #geom_hline(yintercept = 0)+
  ggtitle("Mock Subtracted Analyte Concentrations")

```

This is just donors 317,318,323 and 324, which were interesting in the microarray

```{r subset scatter plots, fig.width = 6,fig.height= 13, echo=FALSE}

#Make plots just showing data for donors 317, 318, 323,324

subset<-mergeMock%>%
  filter(Donor == "317"|Donor == "318"|Donor =="323"|Donor =="324")

#plot it

ggplot(subset, aes(x = Timepoint, y = logMockSubtracted))+
   geom_point(aes(color = Donor),position = position_swarm(0.1), alpha = 0.75, size = 3)+
   labs( y = "log10 Estimated Stock Concentration pg/ml")+
   facet_grid(Assay~Virus, scales = "free_x")+
   theme(axis.text.x = element_text(size = 11))+
  ggtitle("Mock Subtracted Analyte Concentrations")
```

```{r setup for individual analyte HEATMAP, echo=FALSE}

#I want to make a heatmap that has the following data

#NO 8hr timepoint data

#tiles represent foldchange of Mock subtracted 24hr data over mock-subtracted 3hr. So if 24hr>3hr, you get a ratio <1 and if 24hr <3hr you get a ratio >1.

#recall that log(a)-log(b) = log(a/b) so I'm going to do subtration, not division to get the log ratio FC.

#Want gradient intensity to max out at FC of 10.

#Exclude 8hr data.

no8mergeMock <- mergeMock %>%
  filter(Timepoint != "8")

# Extract the 3hr data and select a few columns

timepoint3<- no8mergeMock %>%
  filter(Timepoint == "3")%>%
  select(Donor, Assay, Sample, Virus, logMockSubtracted)

#extract 24 hr data and select a few columns
timepoint24<- no8mergeMock %>%
  filter(Timepoint =="24")%>%
  select(Donor, Assay, Sample, Virus, logMockSubtracted)

```

```{r merge 24 and 3, check NAs, echo=FALSE}
#merge 24hr and 3hr data

merge24And3<- merge(timepoint24,timepoint3, by  = c("Donor","Assay","Virus"), all.x = TRUE, all.y = TRUE)

#There are 100 rows where we have data for 24hr but NOT for 3hr.
which24no3<-merge24And3[which(is.na(merge24And3$Sample.y)),]

#there are NO rows where we have data for 3hr but not for 24hr
which3no24<-merge24And3[which(is.na(merge24And3$Sample.x)),]

#remove the NA's for tidyness

merge24And3 <- na.omit(merge24And3)

#Fix up colnames      
names(merge24And3)[4:7]<- c("Sample.24", "logMockSubtracted.24","Sample.3", "logMockSubtracted.3")
```

```{r get ratio, echo=FALSE}

#Do the subtraction to get a foldchange.

merge24And3 <- merge24And3 %>%
  mutate(logMockSub24_minus_logMockSub3 = logMockSubtracted.24-logMockSubtracted.3)
```

```{r setup logFC heatmap, echo=FALSE}

#I want the scale to have max intensity a 10-fold change in either direction (log10(10)=1 on my scale and log10(0.1)= -1 on my scale) so I am going to set all values >=1 as 1 and all values <=-1 as -1.

#set all values >=1 as 1

merge24And3 <- merge24And3 %>%
  mutate(max10fold =
           ifelse(
             logMockSub24_minus_logMockSub3 >=1,
         1,
           ifelse(
             logMockSub24_minus_logMockSub3 <=-1,
         -1,logMockSub24_minus_logMockSub3)
           )
  )

```

```{r logFC heatmap, echo=FALSE}

#Note that we should NOT have data for the following samples  because we didnt have sup for them

# 323-T2-Anything (i.e. no SD90 data)
# Didnt have sup for 326 T1 OR T2 (and this shows T3/T1 so we don't have anything)

ggplot(merge24And3, aes(x = Donor, y = Assay))+
  geom_tile(aes(fill = max10fold), color = "gray")+
  scale_fill_gradient2(low = "green", mid = "white", high = "red", midpoint = 0, name = "log10 Fold Change")+
  facet_wrap(~Virus, scales = "free_x")+
  ggtitle("log fold change of concentrations at 24hrs over 3 hrs")
```

Here is the same heat map with some changes to to Assay names and order:

```{r rename analytes logFC heatmap, echo=FALSE}

#make a new column to hold the new names
merge24And3<-merge24And3 %>%
  mutate(altAssayNames = ifelse(Assay == "MIG",
                                "CXCL9",
                                ifelse(Assay == "IP-10",
                                       "CXCL10",
                                       ifelse(Assay == "IL1-RA","IL-1RA",as.character(Assay)))))

#factor the new assay names in the order FH wanted:
#Note that the FIRST level will be at the bottom of the heat map axis!!!

merge24And3$altAssayNames<-factor(merge24And3$altAssayNames)

merge24And3$altAssayNames<-factor(merge24And3$altAssayNames, levels(merge24And3$altAssayNames)[c(3,14,7,18,16,17,8,6,5,12,19,4,20,15,13,11,9,10,1,2)])

ggplot(merge24And3, aes(x = Donor, y = altAssayNames))+
  geom_tile(aes(fill = max10fold), color = "gray")+
  scale_fill_gradient2(low = "green", mid = "white", high = "red", midpoint = 0, name = "log10 Fold Change")+
  facet_wrap(~Virus, scales = "free_x")+
  labs( y = "Assay")+
  ggtitle("log fold change of concentrations at 24hrs over 3 hrs")